Mergeable Types

### 2017-02-27 paper/examples.tex
edits
Evaluation

\section{Mergeable Types}

@@

The \drawsome example demonstrates the utility of the \name
programming model in building concurrent applications with conceptual
state sharing. The model lets concurrent computations be composed
around any mergeable type. We now present a series of examples that
demonstrate how mergeable types are built: either by defining merge
functions for OCaml data types, or by composing existing mergeable
types.

@@

As an abstract data type, lists support three
operations: \C{insert x i l}, to insert an element \C{x} at position
\C{i} in the list \C{l}, \C{delete i l}, to delete the element at
\C{i}'th position, and \C{subst i x l}, to substitute the element at
the \C{i}'th position with \C{x}.  Lists of mergeable items are
mergeable if they preserve the following properties:
\begin{itemize}
  \item Elements present in both the merging lists will be present in
  the final list. No element is deleted unless it is deleted in at
  least one list.
  \item An element deleted in one list will not reappear in the merged
  list (unlike Amazon's replicated shopping cart~\cite{dynamo}).
  \item The order of elements is retained. That is, if $x$ occurs
  before $y$ in one of the lists, and if both elements are present in
  the merged list, then $x$ occurs before $y$ in the merged list.
\item If an element $y$ is substituted for an element $x$ in one list,
  then $y$ is also substituted for $x$ in the merged list (i.e., $x$
  is absent and $y$ is present), unless $x$ is deleted or substituted
  in the other list. In the former case, deletion wins, and neither
  $x$ nor $y$ occur in the merged list. In the latter case, if $x$ is
  substituted by a different element $z$ in the other list, then the
  merged list substitutes $x$ by a merge of $y$ and $z$, defined as
  per the merge semantics of the list element type.
\end{itemize}

>>>>

\section{Mergeable Types}
\label{sec:merge-examples}

@@

The \drawsome example demonstrates the utility of the \name
programming model in building concurrent applications with conceptual
state sharing that make them suitable for transparent deployment in a
distributed setting. The model lets concurrent computations to be
composed around any mergeable type. We now present a series of
examples that demonstrate how mergeable types are built: either by
defining merge functions for OCaml data types, or by composing
existing mergeable types.

@@

As an abstract data type, lists support three
operations: \C{insert x i l}, to insert an element \C{x} at position
\C{i} in the list \C{l}, \C{delete i l}, to delete the element at
\C{i}'th position, and \C{subst i x l}, to substitute the element at
the \C{i}'th position with \C{x}.  Lists of mergeable items are
mergeable if they preserve the following properties:
\begin{itemize}
  \item Elements present in both the merging lists will be present in
  the final list. No element is deleted unless it is deleted in at
  least one list.
  \item An element deleted in one list will not reappear in the merged
  list (unlike Amazon's replicated shopping cart~\cite{Dynamo}).
  \item The order of elements is retained. That is, if $x$ occurs
  before $y$ in one of the lists, and if both elements are present in
  the merged list, then $x$ occurs before $y$ in the merged list.
\item If an element $y$ is substituted for an element $x$ in one list,
  then $y$ is also substituted for $x$ in the merged list (i.e., $x$
  is absent and $y$ is present), unless $x$ is deleted or substituted
  in the other list. In the former case, deletion wins, and neither
  $x$ nor $y$ occur in the merged list. In the latter case, if $x$ is
  substituted by a different element $z$ in the other list, then the
  merged list substitutes $x$ by a merge of $y$ and $z$, defined as
  per the merge semantics of the list element type.
\end{itemize}

### 2017-02-27 paper/intro.tex
edits

Framing replication as merging leads to a counter implementation that
bears strong similarity to the original sequential one:
  \begin{ocaml}
    module Replicated_Counter = struct
      include Counter
      type mergeable_t = t 
      let merge lca v1 v2 =
         lca + (v1 - lca) + (v2 - lca)
    end
  \end{ocaml}

The role of \C{lca} (least common ancestor) here captures salient
history - the state resulting from the merge of two versions derived
from the same ancestor state should not unwittingly duplicate the
contributions of the ancestor.  This interpretation of a replicated
datatype is thus given in terms of the evolution of a program state
impicitly associated with the different nodes that comprise a
distributed application with merge operations serving to communicate
and reconcile different local states.

@@

>>>>

Framing replication as merging leads to a counter implementation that
bears strong similarity to the original sequential one:
  \begin{ocaml}
    module Replicated_Counter = struct
      include Counter
      let merge lca v1 v2 =
         lca + (v1 - lca) + (v2 - lca)
    end
  \end{ocaml}
The role of \C{lca} (least common ancestor) here captures salient
history - the state resulting from the merge of two versions derived
from the same ancestor state should not unwittingly duplicate the
contributions of the ancestor.  This interpretation of a replicated
datatype is thus given in terms of the evolution of a program state
impicitly associated with the different nodes that comprise a
distributed application with merge operations serving to communicate
and reconcile different local states.

@@

The remainder of the paper is organized as follows.  We present the
\name\ programming model informally in the next section.
Sec.~\ref{sec:merge-examples} develops a series of examples that
illustrate how merge operations can be defined and used.
Sec.~\ref{sec:formalization} defines an operational semantics, and
formalizes our intution of a mergeable type in terms of morphisms over
datatype operations.  We also present soundness (consistency) and
progress guarantees enjoyed by the
semantics. Sec.~\ref{sec:system-model} considers an instantiation of
the model suitable for distributed environments with network
partitions and failures.  Experimental results

, along with
a series of examples illustrating how merge operations can be defined
and used.  Sec.~\ref{sec:
